### 2019-09-15 ic-journal/unique.tex
Final fixes; disabled highlights

\hlD{Now we recall the ``unique solution of
contractions'' theorem}~\cite{sangiorgi2017equations},
\hlD{which weakens} the requirements of Milner's result
(Theorem~\ref{t:Mil89}).

@@

We have $\ct{\til P} \mcontrBIS \ctpp{\til P}$, and 
$\ct{\til Q} \mcontrBIS \ctpp{\til Q}$, 
 by the substitutivity  properties of $\mcontrBIS$ (we exploit here
 the syntactic constraints on sums). Moreover,
 since each hole of the  context $\qctpp$ is underneath at least $n$
 prefixes, applying the definition
 of $ \mcontrBIS$ on the transition 
 $\ct{\til P}\Arr{\mu}  R$, we infer the existence
 of $\qctp$ such that 
$\ctpp{\til P}\Arcap{\mu} \ctp{\til P} \mexpaBIS R$
and 
$\ctpp{\til Q}\Arcap{\mu}  \ctp{\til Q}$.
Finally, again applying the definition of $\mcontrBIS$ on 
$\ct{\til Q} \mcontrBIS \ctpp{\til Q}$, 
we derive 
$\ct{\til Q}\Arcap{\mu} R' \wb \ctp{\til Q}$ \hl{for some $R'$}.
\end{proof}

@@

\begin{proof}
\hlD{We  prove $P_i \wb Q_i$ (for each $i \in I$) by considering} the following relation
\begin{equation*}
\R \DSdefi \{(R,S) \st R \wb \ct{\til P}, S \wb \ct{\til Q} \mbox{~for some context
$\qct$ (\hl{with only guarded sums})} \} \enspace.
\end{equation*}
\finish{below: what is $\equiv$?}

\hl{Obviously we have $(P_i,Q_i) \in \R$ (by taking $C \equiv E_i$, and
the fact that $\mcontrBIS$ implies $\wb$). It remains to}
show that $\R$ is a bisimulation. Suppose \hl{$(R, S) \in \R$ via} a context
$C$. \hl{For any $R'$ such that $R \arr{\mu} R'$, we have to find a} $S'$ with $S \Arcap{\mu}
S'$ and $(R', S') \in \R$. From $R \wb C[{\til P}]$, we derive $C[{\til P}]
\Arcap{\mu} R'' \wb R'$ for some $R''$. Then by Lemma~\ref{l:uptocon},
there exists $C'$ with $R'' \mcontrBIS C'[{\til P}]$ and $C[{\til Q}]
\Arcap{\mu} S'' \wb C'[{\til Q}]$ for some $S''$. \hl{From $S \wb C[{\til
  Q}]$ and $C[{\til Q}] \Arcap{\mu} S''$, by induction and definition of $\wb$, we find
 $S'$} with $S \Arcap{\mu} S'$. This \hl{completes} the
proof, as we have $R' \wb C'[{\til P}]$ and $S' \wb C'[{\til Q}]$ \hl{by
transitivity of $\wb$ and the fact that $\mcontrBIS$ implies $\wb$.
(The other side from $S$ follows in the same manner.)}
\hlD{See} Fig.~\ref{fig:310} \hl{for a visual illustration.}
\end{proof}

@@

Theorem~\ref{t:contraBisimulationU} brings a new proof technique for
bisimilarity, which is much less restrictive than Milner's
Theorem~\ref{t:Mil89} (but with the additional costs of checking $\mcontrBIS$
in addition to $\wb$).
However, comparing with Milner's Theorem~\ref{t:Mil89s1}, \hlD{there is
a constraint: the involved CCS expression may only
use guarded sums. The is mainly due to the fact that $\wb$ is not a
 congruence and also $\mcontrBIS$ is not a  precongruence.
Inspired by rooted bisimilarity, to  eliminate the restriction
on guarded sums, we have extended the idea of contractions by moving
to \emph{rooted contractions}:}

@@

\hl{The above definition was found with the help of interactive theorem proving.}
\hlD{The following two principles were adopted when searching
for a possible definition: (1) the definition should not be coinductive,
along the lines of rooted bisimilarity} $\rapprox$ (Def.~\ref{d:rootedBisimilarity});
(2) the definition should be built on top of  the existing \emph{contraction}
relation $\mcontrBIS$.
% \hl{Then several candidates were quickly checked (using very similar
%   proof steps): the key is to prove its transitivity, and whenever the
%   proof attempts failed we tried to fix the definition. Eventually the
%   above definition was found and shown to be a precongruence, i.e. it
%   is preserved by all CCS operators except the Recursion, same as in
%   the cases of $\sim$ and $\rapprox$.}
\hlD{Furthermore, we needed to show that the definition found is the
best possible one, by showing that it yields the coarsest precongruence
contains in $\mcontrBIS$; the proof is similar
to  the analogous result for $\rapprox$}. \hlD{See} Section~\ref{s:coarsest} for
more details.

@@

\hl{The next lemma is actually Lemma 3.13} of~\citep[p.~102]{Mil89}, \hlD{and
is needed} to prove Milner's ``unique solution of equations for
$\sim$'' (Theorem~\ref{t:Mil89s1}):
\begin{lemma}
  \label{lem:milner313}
  If the variables ${\til X}$ are weakly guarded in $E$, and $E\{\til
  P/\til X\} \arr{\alpha} P'$, then $P'$ takes the form $E'\{\til
  P/\til X\}$ (for some expression $E'$), and moreover, for any ${\til Q}$,
  $E\{\til Q/\til X\} \arr{\alpha} E'\{\til Q/\til X\}$.
\end{lemma}

@@

Following the same steps in the proof of
Theorem~\ref{t:contraBisimulationU}
(using Lemma~\ref{l:ruptocon}
in place of Lemma~\ref{l:uptocon}), \hl{we can prove that $(P_i, Q_i)
\in\R$ and $\R$ is indeed a bisimulation. But this only shows $P_i \wb Q_i$.}
%
To further show $P_i \rapprox Q_i$, we appeal to
Lemma~\ref{l:obsCongrByWeakBisim}: \hl{for any $P'$ such that $P_i \arr{\mu} P'$, we have to
  find a $Q'$ with $Q_i \Arr{\mu} Q'$ and $(R',S') \in \R$.
%
From $P_i \rcontr E_i[{\til P}]$, by definition of $\rcontr$ we derive $E_i[{\til P}] \arr{\mu}
P''$ for some $P''$.} Then by Lemma~\ref{lem:milner313} there exists a context
$E'$ with $P'' = E'[{\til P}]$ and $E_i[{\til Q}] \arr{\mu} E'[{\til Q}]$.
%
From $Q_i \rcontr E_i[{\til Q}]$ and $E_i[{\til Q}] \arr{\mu} E'[{\til
  Q}]$, by definition of $\rcontr$ we have $Q_i \Arr{\mu} Q'$ for some
$Q'$ and $Q' \wb E'[{\til Q}]$. This \hl{completes} the proof, as we have
$P' \wb C[{\til P}]$ (by the fact that $\mcontrBIS$ implies $\wb$) and
$Q' \wb C[{\til Q}]$. (The other side from $Q_i$ follows in the same manner.)
\hlD{See} Fig.~\ref{fig:314}.
\end{proof}

>>>>

\hl{Now we recall the ``unique solution of
contractions'' theorem}~\cite{sangiorgi2017equations},
\hl{which weakens} the requirements of Milner's result
(Theorem~\ref{t:Mil89}).

@@

We have $\ct{\til P} \mcontrBIS \ctpp{\til P}$, and 
$\ct{\til Q} \mcontrBIS \ctpp{\til Q}$, 
 from the precongruence property of $\mcontrBIS$ (we exploit here
 the syntactic constraints on sums). Moreover,
 since each hole of the  context $\qctpp$ is underneath at least $n$
 prefixes, applying the definition
 of $ \mcontrBIS$ on the transition 
 $\ct{\til P}\Arr{\mu}  R$, we infer the existence
 of $\qctp$ such that 
$\ctpp{\til P}\Arcap{\mu} \ctp{\til P} \mexpaBIS R$
and 
$\ctpp{\til Q}\Arcap{\mu}  \ctp{\til Q}$.
Finally, again applying the definition of $\mcontrBIS$ on 
$\ct{\til Q} \mcontrBIS \ctpp{\til Q}$, 
we derive 
$\ct{\til Q}\Arcap{\mu} R' \wb \ctp{\til Q}$ \hl{for some $R'$}.
\end{proof}

@@

\begin{proof}
\hl{We  prove $P_i \wb Q_i$ (for each $i \in I$) by considering} the following relation
\begin{equation*}
\R \DSdefi \{(R,S) \st R \wb \ct{\til P}, S \wb \ct{\til Q} \mbox{~for some context
$\qct$ (\hl{with only guarded sums})} \} \enspace.
\end{equation*}

Obviously we have $(P_i,Q_i) \in \R$ (by taking $C = E_i$, and
the fact that $\mcontrBIS$ implies $\wb$). It remains to
show that $\R$ is a bisimulation. Suppose \hl{$(R, S) \in \R$ via} a context
$C$. \hl{For any $R'$ such that $R \arr{\mu} R'$, we have to find a} $S'$ with $S \Arcap{\mu}
S'$ and $(R', S') \in \R$. From $R \wb C[{\til P}]$, we derive $C[{\til P}]
\Arcap{\mu} R'' \wb R'$ for some $R''$. Then by Lemma~\ref{l:uptocon},
there exists $C'$ with $R'' \mcontrBIS C'[{\til P}]$ and $C[{\til Q}]
\Arcap{\mu} S'' \wb C'[{\til Q}]$ for some $S''$. \hl{From $S \wb C[{\til
  Q}]$ and $C[{\til Q}] \Arcap{\mu} S''$, by induction and definition of $\wb$, we find
 $S'$} with $S \Arcap{\mu} S'$. This \hl{completes} the
proof, as we have $R' \wb C'[{\til P}]$ and $S' \wb C'[{\til Q}]$ \hl{by
transitivity of $\wb$ and the fact that $\mcontrBIS$ implies $\wb$.
(The other side from $S$ follows in the same manner.)}
See Fig.~\ref{fig:310} for a visual illustration.
\end{proof}

@@

Theorem~\ref{t:contraBisimulationU} brings a new proof technique for
bisimilarity, which is much less restrictive than Milner's
Theorem~\ref{t:Mil89} (but with the additional costs of checking $\mcontrBIS$
in addition to $\wb$).
However, comparing with Milner's Theorem~\ref{t:Mil89s1}, \hl{there is
a constraint: the involved CCS expression may only
use guarded sums. The is mainly due to the fact that $\wb$ is not a
 congruence and also $\mcontrBIS$ is not a  precongruence.
Inspired by rooted bisimilarity, to  eliminate the restriction
on guarded sums, we have extended the idea of contractions by moving
to \emph{rooted contractions}:}

@@

\hl{The above definition was found with the help of interactive theorem proving.}
\hl{The following two principles were adopted when searching
for a possible definition: (1) the definition should not be coinductive,
along the lines of rooted bisimilarity} $\rapprox$ (Def.~\ref{d:rootedBisimilarity});
(2) the definition should be built on top of  the existing \emph{contraction}
relation $\mcontrBIS$.
% \hl{Then several candidates were quickly checked (using very similar
%   proof steps): the key is to prove its transitivity, and whenever the
%   proof attempts failed we tried to fix the definition. Eventually the
%   above definition was found and shown to be a precongruence, i.e. it
%   is preserved by all CCS operators except the Recursion, same as in
%   the cases of $\sim$ and $\rapprox$.}
\hl{Furthermore, we needed to show that the definition found is the
best possible one, by showing that it yields the coarsest precongruence
contains in $\mcontrBIS$; the proof is similar
to  the analogous result for $\rapprox$}. \hl{See} Section~\ref{s:coarsest} for
more details.

@@

The next lemma is actually Lemma 3.13 of~\citep[p.~102]{Mil89}, and
is needed to prove Milner's ``unique solution of equations for
$\sim$'' (Theorem~\ref{t:Mil89s1}):
\begin{lemma}
  \label{lem:milner313}
  If the variables ${\til X}$ are weakly guarded in $E$, and $E\{\til
  P/\til X\} \arr{\alpha} P'$, then $P'$ takes the form $E'\{\til
  P/\til X\}$ (for some expression $E'$), and moreover, for any ${\til Q}$,
  $E\{\til Q/\til X\} \arr{\alpha} E'\{\til Q/\til X\}$.
\end{lemma}

@@

Following the same steps in the proof of
Theorem~\ref{t:contraBisimulationU}
(using Lemma~\ref{l:ruptocon}
in place of Lemma~\ref{l:uptocon}), \hl{we can prove that $(P_i, Q_i)
\in\R$ and $\R$ is indeed a bisimulation. But this only shows $P_i \wb Q_i$.}
%
To further show $P_i \rapprox Q_i$, we appeal to
Lemma~\ref{l:obsCongrByWeakBisim}: \hl{for any $P'$ such that $P_i \arr{\mu} P'$, we have to
  find a $Q'$ with $Q_i \Arr{\mu} Q'$ and $(R',S') \in \R$.
%
From $P_i \rcontr E_i[{\til P}]$, by definition of $\rcontr$ we derive $E_i[{\til P}] \arr{\mu}
P''$ for some $P''$.} Then by Lemma~\ref{lem:milner313} there exists a context
$E'$ with $P'' = E'[{\til P}]$ and $E_i[{\til Q}] \arr{\mu} E'[{\til Q}]$.
%
From $Q_i \rcontr E_i[{\til Q}]$ and $E_i[{\til Q}] \arr{\mu} E'[{\til
  Q}]$, by definition of $\rcontr$ we have $Q_i \Arr{\mu} Q'$ for some
$Q'$ and $Q' \wb E'[{\til Q}]$. This \hl{completes} the proof, as we have
$P' \wb C[{\til P}]$ (by the fact that $\mcontrBIS$ implies $\wb$) and
$Q' \wb C[{\til Q}]$. (The other side from $Q_i$ follows in the same manner.)
See Fig.~\ref{fig:314} for a visual illustration.
\end{proof}
